---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Clement LEFEVRE 10th April 2017.

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(statsr)
library(ggcorrplot)
library(gridExtra)
library(grid)

source('multiplot.r')

```

### Load data


```{r load-data}
load("gss.Rdata")
```



* * *

## Part 1: Data

This dataset is a simplified sample of the official General Social Survey (GSS).

The [**GSS**](http://gss.norc.org/documents/codebook/GSS_Codebook_intro.pdf) is a survey focusing on US society, gathering informations related to **societal behaviors and attitudes**.

According to Wikpedia, "*It is one of the most influential studies in the social sciences, and is frequently referenced in leading publications, including The New York Times, The Wall Street Journal, and the Associated Press.*"

It has been used in 25,000 scholarly publications and every year by 400.000 students.

This study has been done for 40 years.

The data were gathered via face-to-face interviews by US households, for an average of 90 minutes.

#### Sampling methodology

**Random sampling** is implemented via strata : geographical area, sex, race, etc.



#### Case

Cases here are **adults living the US**.

#### Scope of inference

From the elements exposed (random sampling, no treatment assigned to random subjects), we can state that this is an **observational** study, from which we can extract **associations** and considere the model as **generalizable**,  but no we cannot conclude on any **causality** between the explanatory variables (demographics and social) and  the response (opinion, attitude). 


* * *

## Part 2: Research question

As a french, two factors in the US society seem very specifics to me : the inherent racial dichotomy and the attitude towards guns.

Based on this dataset, i'd like to evaluate the possible **association** between **race** and **gun ownership**, and its **evolution** over years.


This short analysis will try to evaluate whether both are somehow related.


* * *

## Part 3: Exploratory data analysis


First, we can have a look at the distribution of interviewees's race over the years  :

```{r}
group_per_year <- gss %>% group_by(year) %>% summarise(count=n())  
average_interviews_per_year <-mean(group_per_year $count)

group_per_year <- gss %>% group_by(year,race) %>% summarise(total_interviews = n())
ggplot(data=group_per_year,aes(x=year,y=total_interviews,fill=race)) + geom_bar(stat='identity')+geom_hline(yintercept = mean(average_interviews_per_year),colour="blue", linetype="dashed")+annotate("text", x=1980, y=average_interviews_per_year*1.1, label="average interviews", color = "blue")
summary(group_per_year)
```


Three elements to be noticed here :

- the average yearly **number of interviewees** is relatively **stable** over year.
- frequency of interviews switched from every year to every two years from 1993 onwards.
- **race split** with the exception of 2 years (1982,1987 : black oversampling) is also **stable.**


Now we can focus on race and gun ownership.
We first create a dataframe with the variables of interest and some additional ones :

- `year` : year of interview
- `age` : Age of respondent
- `sex` : Respondent's sex
- `fear`: Afraid to walk at night in neighborhood
- `owngun` : Have gun in home
- `educ` :  Highest year of school completed
- `coninc` : Total family income in constant dollars

```{r}
weapon_df <- gss %>% select(year,fear,owngun,sex,age,race,educ,coninc)
str(weapon_df)
```



Let's see for both variable the non-answered ratio :
```{r}
as.data.frame(colMeans(is.na(weapon_df))*100)
```

From this table, we can say that even if 40% of interviewees did not provide an answer for `fear` and `owngun`, it seems, according to the sample size (57061 interviewees), somehow acceptable.

We can now remove the N/A from the dataframe :

```{r}
nrow(weapon_df)
weapon_df<- weapon_df %>% na.omit()
nrow(weapon_df)


```


Now we can quickly see the Pearson correlation between those variables.
First, we convert categorical variables into numerics values :


```{r}
weapon_df<- weapon_df %>% mutate(fear_num=ifelse(fear=='Yes',1,0))
weapon_df<- weapon_df %>% mutate(owngun_num=ifelse(owngun=='Yes',2,ifelse(owngun=='Refused',1,0)))

weapon_df<- weapon_df %>% mutate(sex_num=ifelse(sex=='Male',1,0))
weapon_df<- weapon_df %>% mutate(race_num=ifelse(race=='Black',0,ifelse(race=='White',2,1)))
                            
```


Let's plot a correlogramm :

```{r}

weapon_df_numeric<- weapon_df %>% select(-c(fear,sex,race,owngun))

corr <- round(cor(weapon_df_numeric),3)
ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of Selected variables", 
           ggtheme=theme_bw)
```


Well, some elements here :

- **education level** is strongly correlated to the **level of incomes**.(*not relevant for our question*)
- the **fear** factor is correlated to **gender** : men fear less walking in street at night.
- for **gun ownership**, the most correlated variables are in order : 
    1. **race** (whites are more likely to own guns than Others and blacks, $\rho$ = 0.15),
    2. **sex** (men are more likely to own guns),
    3. **income** (the more income, the more frequent is gunownership).
  


Now we can prepare a dataframe with gun ownership grouped by year and race for both women and men : 
```{r}

groupy_gun_race_male<- weapon_df %>% filter(sex=='Male') %>% group_by(year,race,owngun) %>% summarise(total=n(), median_age=median(age)) %>% mutate(ratio=total/sum(total)) %>% filter(owngun=='Yes')

groupy_gun_race_female<- weapon_df  %>% filter(sex=='Female')  %>% group_by(year,race,owngun) %>% summarise(total=n(), median_age=median(age)) %>% mutate(ratio=total/sum(total)) %>% filter(owngun=='Yes')

```

And we plot it :

```{r}
p0<-ggplot(groupy_gun_race_male,aes(x=year,y=ratio*100,col=race)) + geom_line() + theme(axis.title.x=element_blank(),axis.title.y=element_blank()) +labs(title = "Male gun ownership %") +scale_y_continuous(limits = c(0, 60))

p1<-ggplot(groupy_gun_race_female,aes(x=year,y=ratio*100,col=race)) + geom_line() + theme(axis.title.x=element_blank(),axis.title.y=element_blank()) +labs(title = "Female gun ownership %")+scale_y_continuous(limits = c(0, 60))

p2<-ggplot(groupy_gun_race_male,aes(x=year,y=median_age,fill=race,color=race))+geom_line()+ theme(axis.title.x=element_blank(),axis.title.y=element_blank()) +labs(title = "Male gun owners median age")+scale_y_continuous(limits = c(20, 65))

p3<-ggplot(groupy_gun_race_female,aes(x=year,y=median_age,fill=race,color=race))+geom_line() + theme(axis.title.x=element_blank(),axis.title.y=element_blank()) +labs(title = "Female  gun owners median age")+scale_y_continuous(limits = c(20, 65))



grid_arrange_shared_legend(p0, p2, p1, p3)
  
```


A few comments :

- Overall gun ownership decreased over the year.
- Overall median age increased.
- Men of all races have higher gun ownership rate than women.
- The gap between whites and blacks gun ownership increased.
  
  
__Out of the current scope__, i'd like to see the incomes level relationship with gun ownership :
Please take note that I remove the 0.5% of top incomes that i considere, maybe wrongly, as outliers.

```{r}

p0<- ggplot(weapon_df%>% filter(year==1973)  %>% filter(coninc<quantile(coninc,.995)), aes(x=owngun, y=coninc,fill=owngun)) + geom_boxplot()+scale_y_continuous(limits = c(0, 90000))+ theme(legend.position="none") +labs(title= 'Income distribution - 1973')

p1<- ggplot(weapon_df%>% filter(year==2012)  %>% filter(coninc<quantile(coninc,.995)), aes(x=owngun, y=coninc,fill=owngun)) + geom_boxplot()+scale_y_continuous(limits = c(0, 90000))+ theme(legend.position="none")+labs(title = 'Income distribution - 2012')

multiplot(p0, p1,cols = 2)
```


We can clearly observe that during a 40 years time period, the gun ownership is still related to the income level. Rich households are more prone to won a gun, but compared to 1973, gun owners in 2012 are less wealthy. 

In the next part, we will proceed to the simple hypothese testing, whether the mean income level amongst gun ownership categories is similar.


* * *

## Part 4: Inference



### Gun ownership rate evolution over time

Before we start testing the independance of gun ownership with race, we can as a starter check whether overall gun ownership did significantly changed over a period of 40 years.

We can proceed by comparing the gun ownership ration in 1972 with 2012 with the following hypotheses :

$H_{0}:ownership_{1972}-ownership_{2012}=0$

$H_{A}:ownership_{1972}- ownership_{2012}\ne 0$

We are in the context of one numerical variable (**gun ownership ratio**) and one categorical (i considere the **year** as category). 


We can compute the corresponding proportions :

```{r}
ownership1973 <- weapon_df %>% filter(year==1973) %>%group_by(owngun)  %>% summarise(count=n())%>% mutate(ratio=count/sum(count), total = sum(count)) %>% filter(owngun=='Yes')
ownership1973

ownership2012 <- weapon_df %>% filter(year==2012) %>%group_by(owngun)  %>% summarise(count=n())%>% mutate(ratio=count/sum(count), total = sum(count))%>% filter(owngun=='Yes')
ownership2012

```


Now we can compute the pooled proportion of gun owners :

```{r}
p_pool <- (661+390)/(1372+1146)
p_pool
```

Let's check the success failure conditions :

$n_{1973} \times  \hat{p}_{pool} \ge 10$

$(1-n_{1973}) \times  \hat{p}_{pool} \ge 10$

$n_{2012} \times  \hat{p}_{pool} \ge 10$

$(1-n_{2012}) \times  \hat{p}_{pool} \ge 10$

```{r}
p_pool * 1372
(1-p_pool)*1372
p_pool * 1146
(1-p_pool)*1146
```

So far, so good. We can compute the standard estimate SE :

$SE = \sqrt{\frac{\hat{p}_{pool}(1-\hat{p}_{pool})}{n_{1973}}+ \frac{\hat{p}_{pool}(1-\hat{p}_{pool})}{n_{2012}}}$

```{r}
SE<- sqrt(p_pool*(1-p_pool)/1372 + p_pool*(1-p_pool)/1146)
SE
```

Now we can compute the p_value :

```{r}
Z<- (0.3403141-0.4817784)/SE
Z
2*pnorm(Z)

```
```{r}
ggplot(data.frame(x = c(-8, 8)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = SE)) + geom_vline(xintercept = Z,,col='blue',linetype="dashed")+geom_vline(xintercept = -Z,col='blue',linetype="dashed")  +annotate("text", x=-Z*.95, y=10, label="Z", color = "blue")  +annotate("text", x=Z*.95, y=10, label="-Z", color = "blue")
```


From the Z value, it is obvious that the probability that the gun ownership rate in 2012 is different than the one in 1973 occurs by chance is almost zero. This has been confirmed by the chart above.


This conclusion is confirmed by the following article from the Wapo : [**"ownership fall"**](https://www.washingtonpost.com/news/wonk/wp/2016/06/29/american-gun-ownership-is-now-at-a-30-year-low/?utm_term=.f4605793b86f)


### Gun Ownership amongst races and over time
Now, we can test the independance of the gun ownership rate from the race :

#### Hypotheses

$H_{0}:$ Race and gun ownership are independent.

$H_{A}:$ Race and gun ownership are dependent.



#### Chi-square test of independance 

We want to test the independance of gun ownership ('Yes','No') with race ('White', 'Black','Other').
*Nota Bene :* I removed the "Refusal" category from the sample, as it does not provide evidence of the ownership or not.


We are in the case of two categorical variables, with at least 1 with more than 2 levels, the only testing method i see here is the $\chi^2$ test.

##### Conditions for the chi-square test:
1. Independence: Sampled observations must be independent. 	

    ‣ random sample/assignment	&#10004; *Checked*

    ‣ if sampling without replacement, n < 10% of population	&#10004; *Checked*

    ‣ each case only contributes to one cell in the table	 &#10004; *Checked*

2. Sample size: Each particular scenario (i.e. cell) must have at least 5 expected cases. &#10004; *Checked*


I'll use the R function *chisq.test*, but first the appropriate table ownership/race :

```{r}
gun_race <-weapon_df %>% filter(year==2012)  %>% filter(owngun=='Yes' | owngun=='No',owngun!='Refused')

gun_race <- droplevels(gun_race)
gun_race <- table(gun_race$owngun,gun_race$race)
gun_race
chisq.test(gun_race)
```

For the year 2012, we reject the $H_{0}$  : Race and gun ownership are not independant.

What could be interesting is to see the evolution of this p_value over the years :

We can write a generic function to compute the corresponding p.value :

```{r}
chi2_independance_pvalue<- function(df,year_){
gun_race <-df %>% filter(year==year_)  %>% filter(owngun=='Yes' | owngun=='No')

gun_race <- droplevels(gun_race)
gun_race <- table(gun_race$owngun,gun_race$race)

chi2<- chisq.test(gun_race)

return(chi2$p.value)
}
```


Let's compute the evolution of this p.value along the years :

```{r}
gun_asked<-gss %>% group_by(year,owngun) %>% filter(!is.na(owngun)) 
year_asked<-unique(gun_asked$year)
```

```{r}
# My deep apologies, i could not write a cleaner way to create a dataframe from a loop.
x<-list()
for (year in year_asked){
 chi2 <-  chi2_independance_pvalue(weapon_df,year)
 x[as.character(year)]<-list('pvalue'=chi2)
}

 
df_pvalues<-as.data.frame(x) %>% gather() 

colnames(df_pvalues)<-c('yearX','pvalue')
df_pvalues<- df_pvalues %>% separate(yearX,into = c("X", "year"), sep = "X")

ggplot(df_pvalues ,aes(x=year,y=pvalue)) + geom_point()+ theme(axis.text.x=element_text(angle=45, hjust=1)) +geom_hline(yintercept = .05,col='red',linetype="dashed")  

```


The pvalue, with the exception of the year 1976, are all far below 0.05.

We might exclude this 1976 point as an outliers, and plot the log of the pvalue :

```{r}
df_pvalues$log_pvalue<-log(df_pvalues$pvalue)

ggplot(data=df_pvalues %>% filter(pvalue<.1),aes(x=as.numeric(year),y=log_pvalue)) + geom_point() + 
  geom_smooth(method='lm')
```


What we observe here is a decreasing trend towards zero for the p-value.
I would conclude that over the 40 years, the chance of rejection of the hypothese of independance between gun ownership and race has got stronger.


### Mean income equality test amongst gun ownership response (*out of scope*)

#### Hypotheses

$H_{0}:$  The average income level is the same across all gun ownership categories.

$H_{A}:$  The  average income level differs between at least one pair of gun ownership category.

#### ANOVA (*bites the dust*)

In this last test, i would like to implement an Analyse of variance (ANOVA) for the mean income for different categories of gun ownership response ('Yes','No', 'Refusal').

As seen in the EDA part, for 1973 and 2012, we noticed a gap for the **median** income amonst the gun owners and the others.

Can we prove the **mean** also differ via an Anova test ?

I first compute a linear model for both years :
```{r}

lm.1973 <- lm(coninc ~ owngun, data=weapon_df %>% filter(year==1973))
lm.2012 <- lm(coninc ~ owngun, data=weapon_df %>% filter(year==2012))

summary(lm.1973)
summary(lm.2012)
```

It looks already promising, the Yes/No gun ownership have in both years a very low t-value, meaning they are significant in the linear model. By the way, not having a gun means in this context having also lower incomes.

Now we can compute the F statistic and the corresponding p_value :

```{r}
anova(lm.1973)
anova(lm.2012)
```

For both years, the p_value for the F statistic is well below the significant level, we can reject the null hypothese. The mean income discrepancies amongst gun ownership categories **do not happen by chance**.
