---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `gss`. Delete this note when before you submit 
your work. 

```{r load-data}
load("gss.Rdata")
```



* * *

## Part 1: Data


* * *

## Part 2: Research question



* * *

## Part 3: Exploratory data analysis
```{r}
weapon_df <- gss %>% select(year,fear,owngun,pistol,shotgun,rifle,sex,age,race,educ,wrkstat)

groupy1<- weapon_df %>% na.omit() %>% group_by(year,owngun) %>% summarise(total=n()) %>% mutate(ratio=total/sum(total))

ggplot(groupy1,aes(x=year,y=ratio)) + geom_bar(aes(fill = owngun), stat="identity")


```

```{r}
weapon_df %>% group_by(year,sex) %>% summarise(total=n())
head(weapon_df)
```

* * *

## Part 4: Inference

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.