---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
source('multiplot.r)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data

The data used for the research has been collected by the Centers for Disease Control and Prevention (CDC) and US States for the year 2013. This study aimed at analysing behavioral risk factors and preventive health practices and the related health condition. There are 330 variables and 491775 subjects in the 2013 data file. Thoses variables contains informations related to :

- demographics and social context,

- health care access,

- behavioral features (e.g. tobacco use, physical activity),

- current health condition, 

- pathology screening.


#### Cases
The cases are non-institutionalized adults residing in the US.

#### Sampling methodology

Interviews are either via landline of cell phone.
For landline calls, a stratified sampling process has been used for the selection of the number : 1 stratum for high density numbers; an other for medium density.
On top of this sampling technique, the landlines calls have sampled with one randomly selected adult ofr for each household.

For cell calls, a simple random sampling based on the phone number, for person receiving  90% or more of the calls per cell phone.

#### Scope of inference

Since the study does consist in a telephone survey on randomly selected adults and no treatments is randomly assigned to the subjects, we can considere it as an **observational study**.

Therefore we can conclude on **associations** but not on causality between the explanatory variables (behavior and practices) and the response (health condition).

From those points we can infer that the results observed in this study can be **generalizable** on the non-institutionalized adults residing in the US, but no causality can be drawn.



* * *

## Part 2: Research questions

**Research quesion 1:** Is there an association between the quantity of sleep `sleptim1`, tobacco use frequency `smokday2` and alcohol consumtion `avedrnk2` ?

**Research quesion 2:** Are veterans `veteran3` more likely to be single `marital` and more prone to be in poor health `poorhlth` ?

**Research quesion 3:** Are people eating more fruit `fruit1` more likely to use their seatbelt `seatbelt` ?


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research question 1:**
Is there an association between the quantity of sleep `sleptim1`, tobacco use frequency `smokday2` and alcohol consumtion `avedrnk2` ?

```{r}

# useful method to check the NA ratio in the features
NA.ratio <- function(df,cols){
  df_selected <- df %>% select(one_of(cols)) 
  print ('missing answer ratio : ')
  (colMeans(is.na(df_selected))) %>% t()
   
}



features_1 <- c("sleptim1","smokday2","avedrnk2")

NA.ratio(brfss2013,features_1)

df_q1<-brfss2013 %>% select(one_of(features))

dim(df_q1)
dim(df_q1 %>% na.omit())

df_q1_no_na<-df_q1 %>% na.omit()

str(df_q1_no_na)
unique(df_q1_no_na$smokday2)

ggplot(data=df_q1_no_na, aes(x=smokday2,y=sleptim1, fill=smokday2)) + geom_boxplot()
ggplot(data=df_q1_no_na %>% filter(smokday2!='Some days'), aes(x=avedrnk2,y=sleptim1,color=smokday2)) + geom_point(alpha=0.5,position = 'jitter')

```
```{r}
ggplot(data=df_q1_no_na %>% mutate(bin = ntile(avedrnk2, 4)),aes(x=sleptim1,colour=factor(bin))) + geom_histogram()
```



**Research question 2:**
Are veterans `veteran3` more likely to be single `marital` and more prone to be in poor health `poorhlth` ?
```{r}

features_2 <- c("veteran3","marital","poorhlth")

NA.ratio(brfss2013,features_2)

df_q2<-brfss2013 %>% select(one_of(features_2))

dim(df_q2)
dim(df_q2 %>% na.omit())

df_q2_no_na<-df_q2 %>% na.omit()

str(df_q2_no_na)

p1<-ggplot(data=df_q2_no_na %>% filter(veteran3=='No'), aes(x=marital,y=poorhlth,fill=marital)) + geom_boxplot()
p2<-ggplot(data=df_q2_no_na %>% filter(veteran3=='Yes'), aes(x=marital,y=poorhlth,fill=marital)) + geom_boxplot()

 layout <- matrix(c(1,2), nrow = 2, byrow = TRUE)
  multiplot(p1,p2,cols=2,layout=layout)


```



**Research question 3:**

Are people eating more fruit `fruit1` more likely to use their seatbelt `seatbelt` ?

```{r}
features_3 <- c("fruit1","seatbelt")

NA.ratio(brfss2013,features_3)

df_q3<-brfss2013 %>% select(one_of(features_3))

dim(df_q3)
dim(df_q3 %>% na.omit())

df_q3_no_na<-df_q3 %>% na.omit()

str(df_q3_no_na)

ggplot(data=df_q3_no_na, aes(x=seatbelt,y=fruit1)) + geom_boxplot()

ggplot(data=df_q3_no_na, aes(x=fruit1)) + geom_histogram(binwidth = 5)

```

