---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Clement Lefevre 13th April 2017.

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(statsr)
library(GGally)
```

### Load data



```{r load-data}
load("movies.Rdata")
```



* * *

## Part 1: Data

This dataset, as described in the instructions does consist in a **random sample** of **651 US movies** extracted from both **IMDB** and **Rotten Tomatoes** databases.

All have been released between 1970 and 2014. 

Assuming this is an observational study with random sampling, it should be *representative* of US movies. By using random sampling, the resulting prediction model should be *generalizable* to US movies released between 1970 and 2014.




#### Remark regarding generalizability *(out of scope)*

In order to get an idea whether the random sample is generalizable, i compared it with a larger IMDB titles dataset ( *imdb_official_titles.csv* 236,710 different movies) and focused on the **release year distribution**.

I filtered this imdb dataset on the same time period as the movies dataset :

```{r}

min_tht_release_year <- min(movies$thtr_rel_year)
max_tht_release_year<- max(movies$thtr_rel_year)
imdb_titles<-read.csv('imdb_official_titles.csv')
sample_imdb_titles<-sample_n(imdb_titles[(imdb_titles$year>=min_tht_release_year) &
 (imdb_titles$year<=max_tht_release_year),],nrow(movies))

both_datasets<-data.frame(movies$thtr_rel_year,sample_imdb_titles$year)
colnames(both_datasets) <- c("movies","imdb_sample")
both_datasets<-both_datasets %>% gather('dataset','release_year')

ggplot(data=both_datasets,aes(x=release_year,fill=dataset))+geom_histogram(position="identity",alpha=0.3)

```


Both *movies* and *imdb sample* datasets are not normally distributed as shown with the test below :
```{r}
shapiro.test(movies$thtr_rel_year)
shapiro.test(sample_imdb_titles$year)
```
I then  used the Kolmogorvo-Smirnov test to check if both theater release years are equally distributed.

```{r}
ks.test(movies$thtr_rel_year,sample_imdb_titles$year)
```
A qqplot does confirm this result :
```{r}
qqplot(movies$thtr_rel_year,sample_imdb_titles$year)
```

Although the *movies* dataset is a random sample, i cast doubts on the generalizability of such sample.


#### Causality
As we are in a context of **observational studies**, we might conclude some association patterns, but contrary to an experimental study with random samples,  **no causality** that might be deduced from it.



* * *

## Part 2: Research question



* * *

## Part 3: Exploratory data analysis

```{r}

bad_movies <- movies %>% arrange(audience_score) %>% head(200)

head(bad_movies)

plot(bad_movies$audience_score,bad_movies$audience_score)
min(movies$audience_score)
```

```{r}
actors<- grep('actor',colnames(movies),value = TRUE)
dvd<-grep('dvd',colnames(movies),value = TRUE)
tht<-grep('tht',colnames(movies),value = TRUE)
best<-grep('best',colnames(movies),value = TRUE)
critic<-grep('critic',colnames(movies),value = TRUE)
audience<-grep('audience_rating',colnames(movies),value = TRUE)
exclude_cols <-c(c("title","studio","director","imdb_url","rt_url","imdb_rating"),actors,dvd,tht,best,critic,audience)

```


* * *

## Part 4: Modeling


```{r}
movies_simple<- movies %>% select(-one_of(exclude_cols)) %>% na.omit()
lm1<-lm(audience_score~. ,data=movies_simple)

best_lm<-step(lm1,direction ="backward")
summary(best_lm)
```

* * *

## Part 5: Prediction

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 6: Conclusion

